---
title: "Night Sky Report Data Exploration"
author: "Erik Meyer"
date: "`r Sys.Date()`"
output: html_document
params:
  alpha_code:
    label: "Enter 4 letter park code of a new park not in the NPS unit decoder in all caps (UNIT):"
    value: PORE
    input: text
   
 
---

```{r packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


knitr::opts_chunk$set()

packages = c("EnvStats", "reshape2",
             "ggplot2", "ggthemes",
             "pander", "plyr",
             "lubridate", "readxl","tcltk",
             "svDialogs","tcltk2", "tidyverse","vtable",
             "data.table","ggpubr","knitr","readr", "sjmisc","janitor","dplyr",
             "DT","tmap","scales","leaflet", "shiny", "rsconnect", "english",
             "kableExtra","statip","magrittr","flextable","officedown", "officer","forstringr", "geojsonsf","sf","skimr")

## Now load or install & load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)


```

## Data Exploration for Night Sky Reports

Utilize the following table and linked information to inform choosing a site of interest for the data report automation. This should be run first for the reporting park. The ensuing html report can help inform setting parameters for when running the "NSreportmkdown.Rmd". Simply copy and paste the data set id and site name into the parameters. The site you choose can be one of high value night sky resource or one that has been impacted by skyglow. 

```{r dataimport, message=FALSE, include=FALSE}
data<-geojsonsf::geojson_sf("https://www.nps.gov/subjects/nightskies/nsdmap.geojson")
data<-as.data.frame(data)
#create park code field for filtering
data$parkcode<-substr(data$DNIGHT, 1, 4)
#pcode<-"PORE"
pcode<-params$alpha_code
###DATA WRANGLING: NPMAPsQuery dataset
#create report dataset
repdata<-data %>%
  filter(parkcode==pcode)
#Add hmrFull image field
repdata$FULL_LINK<-paste("https://www.nps.gov/subjects/nightskies/graphics/",gsub("\\_.*","_hmrFULL.jpg",repdata$FULLPATH), sep="")
#Add hmrArt image field #subtraction
repdata$ART_LINK<-paste("https://www.nps.gov/subjects/nightskies/graphics/",gsub("\\_.*","_hmrART.jpg",repdata$ARTPATH), sep="")

repdata$Report_Link<-paste("https://www.nps.gov/subjects/nightskies/NSQreport.html?DNIGHT=", repdata$DNIGHT, sep="")

```




Table 1: `r (pcode)` sky brightness metrics derived from the observed images. 




```{r table4, echo=FALSE, ft.align="center", out.width = '100%'}

repdata$VISSTARS_RATIO=as.numeric(repdata$VISSTARS_RATIO)
#create dataset for table4 data
tab4<-repdata %>% select(SITE_NAME,DNIGHT,MID_DATE_LMT, MID_TIME_LMT, HORIZ_MLX, MAXVERT_MLX, ZENITH_LUM_MSA, VISSTARS_RATIO, ALR_POS) %>% 
  arrange(MID_DATE_LMT)%>%
  mutate_if(is.numeric, round, 2) %>%
  mutate(VISSTARS_RATIO=VISSTARS_RATIO*100)

#%>%  #commenting out due to moving away from lost stars variable
                        #mutate(loststars=(1-VISSTARS_RATIO)*100)%>%
                        #relocate(loststars, .after=VISSTARS_RATIO) %>%
                        #select(-VISSTARS_RATIO)

#convert to better date format
tab4$Date<-tab4$MID_DATE_LMT

#need to convert decimal time to hh:mm
tab4$MID_TIME_LMT<-as.character(tab4$MID_TIME_LMT)
#take first part of number and separate out the hours
tab4$Hr<-as.numeric(str_extract_part(tab4$MID_TIME_LMT, before = TRUE, pattern = "."))

#take decimal and convert to minutes
tab4$dec_min<-round((as.numeric(str_extract_part(tab4$MID_TIME_LMT, before = FALSE, pattern = "."))*.01)*60,digits=0)
#Add zero before single digits
tab4$dec_min<-ifelse(nchar(tab4$dec_min)==1,paste("0",tab4$dec_min,sep=""),tab4$dec_min)
#bring both hh and mm together
tab4$hhmm<-paste(tab4$Hr,":",tab4$dec_min,sep = "")


#select data fields for table 4
tab4<-tab4 %>% select(SITE_NAME,DNIGHT,Date, hhmm, HORIZ_MLX, MAXVERT_MLX, ZENITH_LUM_MSA, VISSTARS_RATIO, ALR_POS)  

#change column names for published table
colnames(tab4)<- c("Site Name","Dataset","Date","Time\n (hh:mm)","Horizontal\n Illuminance \n (mlx)", "Max. Vertical \n Illuminance \n (mlx)", "Zenith \n Luminance \n (mag/arcsec2)", "Stars \n Visible \n  (%)", "All-sky Light \n Pollution \n Ratio")

#create dataset for reference values
refdata<- data.frame(SiteName=c('Urban Sky', 'Natural Sky'),dset=c('' , ''),Date=c('' , ''),Time=c('' , ''), hi=c(39.56,0.80), mvi=c(29.04, 0.40), zb=c(18.00, 22.00), ls=c(8,100), alr=c(64.43,0.00) )
#refdata$zb<-(round(refdata$zb, digits=2))

#change column names of reference data set to match tab4 dataset
colnames(refdata)<- c("Site Name","Dataset","Date","Time\n (hh:mm)","Horizontal\n Illuminance \n (mlx)", "Max. Vertical \n Illuminance \n (mlx)", "Zenith \n Luminance \n (mag/arcsec2)", "Stars \n Visible \n  (%)", "All-sky Light \n Pollution \n Ratio")

#combind data table with reference table
tab4<-rbind(tab4,refdata)

#create table formatted for report
ft4<-flextable(tab4) %>% 
  align(j=1,align="left", part="all") %>% 
  align(j=c(2:9),align="center", part="all") %>%
  fontsize(size = 9, part = "header") %>%
  bold(part="header") %>%
  fontsize(size = 9, part = "body") %>%
  padding(padding = 0, part = "all")  %>%
  autofit() %>% 
  border_outer()%>% 
  border_inner()

#find row number for urban sky for relative placement of superscript
urbannum<-which(tab4[1] == 'Urban Sky')
#find row number for natural sky for relative placement of superscript
natnum<-which(tab4[1] == 'Natural Sky')

#Add first superscript (A)
#ft4a<-compose(ft4, i=1, j=4,part="header", value=as_paragraph( "Max. Vertical \n Illuminance",
    #as_sup(" A"),"\n (mlx)"))
#Add second superscript (b)  
ft4a<- compose(ft4, i=urbannum, j=1,part="body", value=as_paragraph( "Urban Sky",as_sup(" A"))) 
#Add third superscript (c)
ft4b<-compose(ft4a, i=natnum, j=1,part="body", value=as_paragraph( "Natural Sky",as_sup(" B")))
#add superscript 2
ft4c<-compose(ft4b, i=1, j=7,part="header", value=as_paragraph( "Zenith \n Luminance \n (mag/arcsec",as_sup("2"),")"))


#print table
ft4c


```


^A^ Rock Creek Park in Washington, DC was used as the urban sky reference.

^B^ The median natural sky is the natural reference condition.







Table 2: `r (pcode)` visual and SQM measurements. 


```{r table5, echo=FALSE, ft.align="center", out.width = '100%'}

#create dataset for table5 data
tab5<-repdata %>% select(SITE_NAME,DNIGHT, MID_DATE_LMT, BORTLE, ZLM, SQM) %>% #Need to verify correct measures
  arrange(MID_DATE_LMT)%>%
  mutate_if(is.numeric, round, 2) 

#convert to better date format
tab5$MID_DATE_LMT<-tab5$MID_DATE_LMT


#change column names for published table
colnames(tab5)<- c("Site Name","Dataset","Date","Bortle\n Class","Limiting\n Magnitude", "SQM \n (mag/arcsec2)")

#create dataset for reference values
refdatabc<- data.frame(SiteName=c('Urban Sky', 'Natural Sky'),dset=c('' , ''),Date=c('' , ''),bc=c(8,1), lm=c(5.2, 7.0), sqm=c(18.36, 22.00))
#refdata$zb<-(round(refdata$zb, digits=2))

#change column names of reference data set to match tab4 dataset
colnames(refdatabc)<-  c("Site Name","Dataset","Date","Bortle\n Class","Limiting\n Magnitude", "SQM \n (mag/arcsec2)")

#combind data table with reference table
tab5<-rbind(tab5,refdatabc)

#create table formatted for report
ft5<-flextable(tab5) %>% 
  align(j=1,align="left", part="all") %>% 
  align(j=c(2:4),align="center", part="all") %>%
  fontsize(size = 9, part = "header") %>%
  bold(part="header") %>%
  fontsize(size = 9, part = "body") %>%
  padding(padding = 0, part = "all")  %>%
  autofit() %>% 
  border_outer()%>% 
  border_inner()

#find row number for urban sky for relative placement of superscript
urbannum<-which(tab5[1] == 'Urban Sky')
#find row number for natural sky for relative placement of superscript
natnum<-which(tab5[1] == 'Natural Sky')

#Add first superscript (A)

ft5a<- compose(ft5, i=urbannum, j=1,part="body", value=as_paragraph( "Urban Sky",as_sup(" A"))) 
#Add third superscript (c)
ft5b<-compose(ft5a, i=natnum, j=1,part="body", value=as_paragraph( "Natural Sky",as_sup(" B")))
#add superscript 2
ft5c<-compose(ft5b, i=1, j=6,part="header", value=as_paragraph( "SQM","\n (mag/arcsec",as_sup("2"),")"))
#print table
ft5c

```



^A^ Rock Creek Park in Washington, DC was used as the urban sky reference.

^B^ The median natural sky is the natural reference condition.


## Compare to entire NPS Dataset
```{r table, echo=FALSE, message=FALSE, warning=FALSE,ft.align="center", out.width = '100%'}
# Create a summary dataframe with selected columns and rounded numeric values
NPS_Night_Sky_Data_ALL <- data %>%
  select(HORIZ_MLX, MAXVERT_MLX, ZENITH_LUM_MSA, VISSTARS_PCT, ALR_POS, BORTLE, ZLM, SQM) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%  # Use across for better readability
  mutate(across(everything(), ~ as.numeric(as.character(.))))  # Ensure all columns are numeric

# Change column names for the published table
colnames(NPS_Night_Sky_Data_ALL) <- c(
  "Horizontal Illuminance (mlx)", 
  "Max. Vertical Illuminance (mlx)", 
  "Zenith Luminance (mag/arcsec)", 
  "Stars Visible (%)", 
  "All-sky Light Pollution Ratio", 
  "Bortle Class", 
  "Limiting Magnitude", 
  "SQM (mag/arcsec²)"
)

# Generate a summary of the dataframe
tab <- skim(NPS_Night_Sky_Data_ALL)
tab   # Print the summary table
```

## Data Exploration Links

```{r appenda, echo=FALSE,results = 'asis'}
#create dataset for table data and fromat on survey date
appenda<-repdata %>% select(DNIGHT,MID_DATE_LMT, SITE_NAME, NARRATIVE,Report_Link,FULL_LINK,ART_LINK) %>% 
  arrange(MID_DATE_LMT) %>%
  mutate(Date = MID_DATE_LMT)



for (i in 1:length(appenda$DNIGHT)) {
            cat(paste("On ", appenda$Date[i], ", at ",appenda$SITE_NAME[i],":", sep=""))
            cat("  \n")
             cat(paste("NP Maps HTML report: ",appenda$Report_Link[i],sep=""))
            cat("  \n")
            cat(paste("Full resolution mosaic: ",appenda$FULL_LINK[i],sep="")) 
            cat("  \n")  
            cat(paste("Estimated artificial sky glow: ",appenda$ART_LINK[i],sep=""))
  cat("  \n") 
  cat("  \n") 
    
}

```



```{r remove, include=FALSE}
rm(list=ls())

```
